FMA: A Dataset For Music Analysis

### 2017-04-18 fma.tex
update

% about 150-200 words.
\begin{abstract}
The Free Music Archive (FMA) is an open and easily accessible dataset that can be used to evaluate several tasks in music information retrieval (MIR) such as music classification and annotation or recommendation.
While content-based MIR tasks have typically been solved with feature engineering,
%with a two-stage approach: engineered features are extracted from music audio signals, and are then used as input to a regressor or classifier.
the community is becoming interested in feature learning and end-to-end learning. %, i.e. learning from signals to labels
That endeavor is however restrained by the limited availability of large audio datasets.
% The primary goal of this dataset is to enable the training of large models, which require lots of training data, which could potentially avert the semantic gap currently observed between the extracted low-level features such as MFCCs and the high-level targets such as genre.
By releasing the FMA, we hope to foster research which will improve the state-of-the-art in several tasks and hopefully surpass the performance ceiling observed in e.g.\ genre recognition (MGR).
% as it happened in computer vision with the release of ImageNet.
The released data is composed of \ntracks tracks, \nartists artists, \nalbums albums, \ngenres genres, for a total of \tduration days of audio and \size GiB, all under permissive Creative Commons licenses.
It features a collection of metadata such as the song title, album, artist name and genres; user data such as play counts and favorites; free-form text such as description, biography and tags; together with full-length and high-quality audio as well as some pre-computed features. We propose a train/validation/test split and three subsets: a small genre-balanced dataset of 8,000 songs from 8 major genres, a medium genre-unbalanced dataset of 25,000 songs from 16 genres, and a 90 GiB version of the full set with clips trimmed to 30s.
This paper describes the dataset and how it was created, proposes some tasks, and evaluates some baselines for MGR.
Code, data, and usage examples are available at \url{https://github.com/mdeff/fma}.

@@

\begin{table*}[t]
	\centering
	\begin{threeparttable}
		\begin{tabular}{l@{\hspace{2.7em}}r
			@{\hspace{2.7em}}>{$}r<{$}
			@{}>{${}={}$}c@{}
			>{$}l<{$}
			@{\hspace{2.7em}}>{$}c<{$}
			@{\hspace{2.7em}}c@{\hspace{2.7em}}c}
		\toprule
		dataset & \#samples & \multicolumn{3}{c}{dimensionality} & \multicolumn{3}{c}{size} \\
		\cmidrule{6-8}
				&           & \multicolumn{3}{c}{}               & \text{scale} & [GiB] & [days] \\
		\midrule
		\href{http://mirg.city.ac.uk/codeapps/the-magnatagatune-dataset}{TagATune} \cite{magnatagatune} &
			25,863 & 29\cdot16\cdot10^3\cdot1 && 4.6\cdot10^5 & 1.2\cdot10^{10} & 3 & 8.7 \\ % MagnaTagATune
		\href{https://research.google.com/audioset/}{AudioSet} \cite{audioset} \tnote{1} &
			2,084,320 & 10\cdot44\cdot10^3\cdot2 && 8.8\cdot10^5 & 1.8\cdot10^{12} & - & 241 \\
		\href{https://labrosa.ee.columbia.edu/millionsong/}{MSD} \cite{msd} \tnote{2} &
			1,000,000 & 49\cdot27\cdot10^3\cdot2 && 2.6\cdot10^6 & 2.6\cdot10^{12} & 625 & 541 \\
		\href{https://github.com/mdeff/fma/}{FMA} \tnote{2} &
			\ntracks & \aduration\cdot44\cdot10^3\cdot2 && 2.4\cdot10^7 & 2.6\cdot10^{12} & \size & \tduration \\
		\midrule
		\href{http://yann.lecun.com/exdb/mnist/}{MNIST} \cite{mnist} &
			70,000 & 28\cdot28\cdot1 && 7.8\cdot10^2 & 5.5\cdot10^{7\phantom0} & 0.011 & - \\
		\href{https://www.cs.toronto.edu/~kriz/cifar.html}{CIFAR-10} \cite{cifar} &
			60,000 & 32\cdot32\cdot3 && 3.0\cdot10^3 & 1.8\cdot10^{8\phantom0} & 0.16 & - \\
		\href{http://www.image-net.org/challenges/LSVRC/2012/}{ILSVRC 2012-17} \cite{imagenet_challenge} \tnote{3,4} &
			1,341,167 & 482\cdot415\cdot3 && 6.0\cdot10^5 & 8.0\cdot10^{11} & 100 & - \\
		\href{http://www.image-net.org}{ImageNet} \cite{imagenet} \tnote{3} &
			14,197,122 & 482\cdot415\cdot3 && 6.0\cdot10^5 & 8.5\cdot10^{12} & 1000 & - \\
		\bottomrule
	\end{tabular}
	\begin{tablenotes}
		\item[1] \href{https://support.google.com/youtube/answer/6039860}{YouTube recommended} sample rate for music videos.
		\item[2] Average duration and sample rate. Most excerpts from \href{https://www.7digital.com}{7digital} are 30 or 60s and 22 or 44kHz \cite{msd_features}.
		\item[3] Average resolutions. Most applications resize or crop to $256\cdot256$ for a scale of $2.6\cdot10^{11}$ and $2.8\cdot10^{12}$.
		\item[4] Object detection / classification task. It is the most used dataset for image classification.
	\end{tablenotes}
	\end{threeparttable}
	\caption{Comparison between the largest public audio and image datasets. Dimensionality is (length $\cdot$ sample rate $\cdot$ \#channels) for audio and (xdim $\cdot$ ydim $\cdot$ \#channels) for images. Scale is the number of samples times the dimensionality. Size is for a (zipped) archive of all \texttt{.mp3} or \texttt{.jpg} in GB, an indication of the quantity of information affected by the size and diversity of the data. The last column is the number of days necessary to listen to the whole available audio.}
	\label{tab:size}
\end{table*}
% Magnatagature dead links: http://tagatune.org/Datasets.html http://tagatune.org/Magnatagatune.html
% Announcement: https://musicmachinery.com/2009/04/01/magnatagatune-a-new-research-data-set-for-mir/

@@

% Value proposition.
\textbf{Large scale.} Large datasets are needed to avoid overtraining and to effectively learn models that incorporate the ambiguities and inconsistencies that one finds with musical categories. They are also more diverse and allows to average out annotation noise.
While FMA features less clips than MSD or AudioSet, every other dataset with readily available quality audio are two orders of magnitude smaller (see \tabref{tab:datasets}).
For statistical learning algorithms, the scale of FMA is comparable to ImageNet and the most used ILSVRC (see \tabref{tab:size}) and shall thus be sufficient for the foreseable future.
Note that while image datasets are larger, the amount of data is similar because of their smaller dimensionality.

@@

\begin{table}
	\centering
	\begin{tabular}{lrrccc}
		\toprule
		dataset & clips & genres & length & \multicolumn{2}{c}{size} \\
		\cmidrule{5-6}
		        &       &        &  [s]   & [GB] & [days] \\
		\midrule
		small  &    8,000 &   8 &  30 & 3.4  & 2.8  \\
		medium &   25,000 &  16 &  30 & 12.2 & 8.7  \\
		large  & \ntracks & 161 &  30 & 90   & 37 \\
		full   & \ntracks & 161 & \aduration & \size & \tduration  \\
		\bottomrule
	\end{tabular}
	\caption{Proposed subsets of the FMA.}
	\label{tab:subsets}
\end{table}

>>>>

% about 150-200 words.
\begin{abstract}
The Free Music Archive (FMA) is an open and easily accessible dataset that can be used to evaluate several tasks in music information retrieval (MIR) such as music classification and annotation or recommendation.
While content-based MIR tasks have typically been solved with feature engineering,
%with a two-stage approach: engineered features are extracted from music audio signals, and are then used as input to a regressor or classifier.
the community is becoming interested in feature learning and end-to-end learning. %, i.e. learning from signals to labels
That endeavor is however restrained by the limited availability of large audio datasets.
% The primary goal of this dataset is to enable the training of large models, which require lots of training data, which could potentially avert the semantic gap currently observed between the extracted low-level features such as MFCCs and the high-level targets such as genre.
By releasing the FMA, we hope to foster research which will improve the state-of-the-art in several tasks and hopefully surpass the performance ceiling observed in e.g.\ genre recognition (MGR).
% as it happened in computer vision with the release of ImageNet.
The released data is composed of \ntracks tracks, \nartists artists, \nalbums albums, \ngenres genres, for a total of \tduration days of audio and \size GiB, all under permissive Creative Commons licenses.
It features a collection of metadata such as the song title, album, artist name and genres; user data such as play counts and favorites; free-form text such as description, biography, tags, and comments; together with full-length and high-quality audio as well as some pre-computed features. We propose a train/validation/test split and three subsets: a small genre-balanced dataset of 8,000 songs from 8 major genres, a medium genre-unbalanced dataset of 25,000 songs from 16 genres, and a 90 GiB version of the full set with clips trimmed to 30s.
This paper describes the dataset and how it was created, proposes some tasks, and evaluates some baselines for MGR.
Code, data, and usage examples are available at \url{https://github.com/mdeff/fma}.

@@

\begin{table*}[t]
	\centering
	\begin{threeparttable}
		\begin{tabular}{l@{\hspace{2.7em}}r
			@{\hspace{2.7em}}>{$}r<{$}
			@{}>{${}={}$}c@{}
			>{$}l<{$}
			@{\hspace{2.7em}}>{$}c<{$}
			@{\hspace{2.7em}}c@{\hspace{2.7em}}c}
		\toprule
		dataset & \#samples & \multicolumn{3}{c}{dimensionality} & \multicolumn{3}{c}{size} \\
		\cmidrule{6-8}
				&           & \multicolumn{3}{c}{}               & \text{volume} & [GiB] & [days] \\
		\midrule
		\href{http://mirg.city.ac.uk/codeapps/the-magnatagatune-dataset}{TagATune} \cite{magnatagatune} &
			25,863 & 29\cdot16\cdot10^3\cdot1 && 4.6\cdot10^5 & 1.2\cdot10^{10} & 3 & 8.7 \\ % MagnaTagATune
		\href{https://research.google.com/audioset/}{AudioSet} \cite{audioset} \tnote{1} &
			2,084,320 & 10\cdot44\cdot10^3\cdot2 && 8.8\cdot10^5 & 1.8\cdot10^{12} & - & 241 \\
		\href{https://labrosa.ee.columbia.edu/millionsong/}{MSD} \cite{msd} \tnote{2} &
			1,000,000 & 49\cdot27\cdot10^3\cdot2 && 2.6\cdot10^6 & 2.6\cdot10^{12} & 625 & 541 \\
		\href{https://github.com/mdeff/fma/}{FMA} \tnote{2} &
			\ntracks & \aduration\cdot44\cdot10^3\cdot2 && 2.4\cdot10^7 & 2.6\cdot10^{12} & \size & \tduration \\
		\midrule
		\href{http://yann.lecun.com/exdb/mnist/}{MNIST} \cite{mnist} &
			70,000 & 28\cdot28\cdot1 && 7.8\cdot10^2 & 5.5\cdot10^{7\phantom0} & 0.011 & - \\
		\href{https://www.cs.toronto.edu/~kriz/cifar.html}{CIFAR-10} \cite{cifar} &
			60,000 & 32\cdot32\cdot3 && 3.0\cdot10^3 & 1.8\cdot10^{8\phantom0} & 0.16 & - \\
		\href{http://www.image-net.org/challenges/LSVRC/2012/}{ILSVRC 2012-17} \cite{imagenet_challenge} \tnote{3,4} &
			1,341,167 & 482\cdot415\cdot3 && 6.0\cdot10^5 & 8.0\cdot10^{11} & 100 & - \\
		\href{http://www.image-net.org}{ImageNet} \cite{imagenet} \tnote{3} &
			14,197,122 & 482\cdot415\cdot3 && 6.0\cdot10^5 & 8.5\cdot10^{12} & 1000 & - \\
		\bottomrule
	\end{tabular}
	\begin{tablenotes}
		\item[1] \href{https://support.google.com/youtube/answer/6039860}{YouTube recommended} sample rate for music videos.
		\item[2] Average duration and sample rate. Most excerpts from \href{https://www.7digital.com}{7digital} are 30 or 60s and 22 or 44kHz \cite{msd_features}.
		\item[3] Average resolutions. Most applications resize or crop to $256\cdot256$ for a volume of $2.6\cdot10^{11}$ and $2.8\cdot10^{12}$.
		\item[4] Object detection / classification task. It is the most used dataset for image classification.
	\end{tablenotes}
	\end{threeparttable}
	\caption{Comparison between the largest public audio and image datasets. Dimensionality is (length $\cdot$ sample rate $\cdot$ \#channels) for audio and (xdim $\cdot$ ydim $\cdot$ \#channels) for images. Volume is the number of samples times the dimensionality. Size is for a (zipped) archive of all \texttt{.mp3} or \texttt{.jpg} in GiB, an indication of the quantity of information affected by the size and diversity of the data. The last column is the number of days necessary to listen to the whole available audio.}
	\label{tab:size}
\end{table*}
% Magnatagature dead links: http://tagatune.org/Datasets.html http://tagatune.org/Magnatagatune.html
% Announcement: https://musicmachinery.com/2009/04/01/magnatagatune-a-new-research-data-set-for-mir/

@@

% Value proposition.
\textbf{Large scale.} Large datasets are needed to avoid overtraining and to effectively learn models that incorporate the ambiguities and inconsistencies that one finds with musical categories. They are also more diverse and allows to average out annotation noise.
While FMA features less clips than MSD or AudioSet, every other dataset with readily available quality audio are two orders of magnitude smaller (\tabref{tab:datasets}).
For statistical learning algorithms, the scale of FMA is comparable to ImageNet and the most used ILSVRC (\tabref{tab:size}) and shall thus be sufficient for the foreseable future.
Note that while image datasets are larger, the amount of data is similar because of their smaller dimensionality.

@@

\begin{table}
	\centering
	\begin{tabular}{lrrccc}
		\toprule
		dataset & clips & genres & length & \multicolumn{2}{c}{size} \\
		\cmidrule{5-6}
		        &       &        &  [s]   & [GiB] & [days] \\
		\midrule
		small  &    8,000 &   8 &  30 & 3.4  & 2.8  \\
		medium &   25,000 &  16 &  30 & 12.2 & 8.7  \\
		large  & \ntracks & 161 &  30 & 90   & 37 \\
		full   & \ntracks & 161 & \aduration & \size & \tduration  \\
		\bottomrule
	\end{tabular}
	\caption{Proposed subsets of the FMA.}
	\label{tab:subsets}
\end{table}
